---
title: "day4"
author: "olivier"
date: "12/4/2021"
output: html_document
---

## Load the data 

I divided it into two parts:  
- the random order of the number
- the bingo matrices

```{r}
random_number <- scan(file = "data/day4", what = "list", nlines = 1, sep = ",") |> 
                    as.numeric()

bingo_raw <- scan(file = "data/day4", what = "list", skip = 1) |> 
                    as.numeric()

bingo_list <- list()

for (i in 1:100){
    start <- (25 * i) - 24
    end <- 25 * i
    bingo_list[[i]] <- matrix(bingo_raw[start:end], ncol = 5, byrow = TRUE)
}
    
```

I think I will replace each matrix number with the time it match a random_number

```{r}
matrix_order <- list()

for (i in 1:100){
matrix_order[[i]] <- matrix(match(bingo_list[[i]], random_number),
                            ncol = 5, byrow = FALSE)
}
# test
#bingo_list[[100]]
#matrix_order[[100]]
```

To complete the lines (row or col) the most important one is the last number to be completed. The lines who get this number at the lowest point win. 

```{r}
row_list <- list()
for (i in 1:100) {
row_list[[i]] <- apply(matrix_order[[i]], 1, max)
}

#row_list[[100]]

which.min(unlist(lapply(row_list, min)))

col_list <- list()

for (i in 1:100) {
col_list[[i]] <- apply(matrix_order[[i]], 2, max)
}

which.min(unlist(lapply(col_list, min)))
```

In my case row win! And the bingo 43 !!!

```{r}

sum((matrix_order[[43]] > 30) * bingo_list[[43]]) * 38

```

## Part 2

We want the last bingo that complete one line. So the max of the min.

```{r}
row_list <- list()
for (i in 1:100) {
row_list[[i]] <- apply(matrix_order[[i]], 1, max)
}

#row_list[[100]]

which.max(unlist(lapply(row_list, min)))

col_list <- list()

for (i in 1:100) {
col_list[[i]] <- apply(matrix_order[[i]], 2, max)
}

which.max(unlist(lapply(col_list, min)))
```


```{r}

sum((matrix_order[[84]] > 86) * bingo_list[[84]]) * 32

apply(matrix_order[[84]], 1, max)

```
