---
title: "day5"
author: "olivier"
date: "12/5/2021"
output: html_document
---

## load data

```{r}
directions <- scan("data/day5", sep = "\n", what =  "character")

library(tidyverse)

directions <- as.data.frame(directions) |> 
                        tidyr::separate(directions, 
                                        into = c("x1", "y1", "x2", "y2"),
                                        convert = TRUE)

```

## Part 1

We need to keep only the horizontal and vertical lines. Find a way to "extend" them. Then sum them to see where more than 1 lines overlap. 

Find a way to "extend" them. -> this was not my solution but David Robinson's one,I need to look into purrr ....

```{r}
axis_directions <- directions[directions$x1 == directions$x2 | directions$y1 == directions$y2 ,]

axis_directions_etendu <- axis_directions %>% 
    dplyr::mutate(long = purrr::map2(x1, x2, seq)) %>% 
    dplyr::mutate(lat = purrr::map2(y1, y2, seq)) %>% 
    tidyr::unnest(long) %>% 
    tidyr::unnest(lat

axis_directions_etendu %>% 
    dplyr::count(long , lat) %>% 
    dplyr::summarize(sum(n > 1))
```

## Part 2

> the lines in your list will only ever be horizontal, vertical, or a diagonal line at exactly 45 degrees.

We have already horizontal and vertical. So we need to "extend" x1 -> x2 and y1 -> y2 then remove horiz and vertical lines to add it to the previous result.

```{r}
diagonal <- directions %>% 
                    dplyr::mutate(long = purrr::map2(x1, x2, seq)) %>% 
                    dplyr::mutate(lat = purrr::map2(y1, y2, seq)) %>% 
                    dplyr::filter(!(x1 == x2 | y1 == y2)) %>% 
                    tidyr::unnest(c(long, lat)) 

rbind(axis_directions_etendu, diagonal) %>% 
    dplyr::count(long, lat) %>% 
    dplyr::summarize(sum(n > 1))
```

